\section{Operational Semantics}
\label{vine:operational}

\input{chap-formalism/operational-semantics-table}

The operational semantics for \bil are shown in
Tables~\ref{bap:taboperstmts} and \ref{bap:taboperexps}.  The abstract
machine configuration is given by the tuple $(\Pi, \Delta, p, s)$
where $\Pi$ is the list of statements, $\Delta$ is the variable
context, $p$ is the statement pointer, and $s$ is the current
statement.  We write $\Delta' = \Delta[x \leftarrow v]$ to indicate
that $\Delta'$ is the same as $\Delta$ except that variable $x$ is
updated with value $v$.  For simplicity, we use $\Delta$ both as a
scalar and a memory context. When ambiguous, such as in the {\sc
  store} rule, we indicate the type of the variable in $\Delta$ in
parentheses. We write $\Pi[p]$ to indicate the statement given by
address $p$.

The operational semantics can be read as follows.  Each step of the
execution is associated with a machine configuration $M = (\Pi,
\Delta, p, s)$.  A transition is given by $M \leadsto M'$ where the
current configuration $M$ matches the left side of $\leadsto$ in the
conclusion (below the horizontal bar), resulting in a state $M'$ to
the right.  The transformation from $M$ to $M'$ is given by the rule
premise (above the horizontal bar). 

{\it Note:} \bap and \bil do not analyze dynamically generated
code. Thus, in a machine state transition $(\Pi, \Delta, p, s)
\rightarrow (\Pi', \Delta, p, s)$, $\Pi = \Pi'$ always. Since $\Pi$
(the list of statements) never changes, we omit $\Pi$ from the rules
for brevity. One could add support for dynamically generated code by
adding rules for updating $\Pi$. 


{\sc assign} and {\sc label} are sequential statements that carry
out the respective operation, then look up and transition to the next
sequential statement $p+1$. The semantics of {\sc label} is a no-op:
we use labels for jump targets.  {\sc assign} updates the variable
context $\Delta$ resulting in a new context $\Delta'$.  As mentioned,
there is no rule for {\tt special}; any program with a {\tt special}
remaining may get stuck.

Control flow is handled by {\tt jmp}%%  (since {\tt cjmp} is a derived
%% form per Chapter~\ref{vine:derived})
%% cjmp is not derived in bap?
. A {\tt jmp} statement
evaluates the jump target $e$ to a value $v$, then looks up the
statement associated with $v$.  The {\sc no-inst} rule terminates
the program in the error state when $v$ is not associated with an
statement. For example, consider the case when a program reads in
user input at location $v$, then issues a jump to $v$.  The user input
will be decoded as statements at run-time.  However, since the
statement comes from user input, we cannot include it in the analysis. In
\bap, we indicate such possibilities by terminating in error.




%\subsubsection{\bap With Functions Operational Semantics}
%\input{vine/operational-semantics-functions-table}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../main"
%%% End: 
